---
title: "Assignment 6: GLMs (Linear Regressios, ANOVA, & t-tests)"
author: "Jun Hu"
date: "Spring 2023"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on generalized linear models. 

## Directions
1. Rename this file `<FirstLast>_A06_GLMs.Rmd` (replacing `<FirstLast>` with your first and last name).
2. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure to **answer the questions** in this assignment document.
5. When you have completed the assignment, **Knit** the text and code into a single PDF file.


## Set up your session 
1. Set up your session. Check your working directory. Load the tidyverse, agricolae and other needed packages. Import the *raw* NTL-LTER raw data file for chemistry/physics (`NTL-LTER_Lake_ChemistryPhysics_Raw.csv`). Set date columns to date objects.

2. Build a ggplot theme and set it as your default theme.

```{r setup2}
#1
library(tidyverse)
library(agricolae)   
library(here)
library(ggplot2)
library(lubridate)
here()

NTL_Chemphy <- read.csv(here("Data/Raw/NTL-LTER_Lake_ChemistryPhysics_Raw.csv"), stringsAsFactors = TRUE)

NTL_Chemphy$sampledate <- as.Date(NTL_Chemphy$sampledate, format = "%m/%d/%y")

#2
mytheme <- theme_classic(base_size = 14) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top")
theme_set(mytheme)

```

## Simple regression
Our first research question is: Does mean lake temperature recorded during July change with depth across all lakes?

3. State the null and alternative hypotheses for this question:
> Answer:
H0: the slope of the lake temperature recorded during July vs. depth across all lakes is equal to zero (i.e., the lake temperature recorded during July will not change with depth across all lakes); Ha: the slope of the lake temperature recorded during July vs. depth across all lakes is not equal to zero (i.e., the lake temperature recorded during July will change with depth across all lakes).


4.  Wrangle your NTL-LTER dataset with a pipe function so that the records meet the following criteria: 
 * Only dates in July. 
 * Only the columns: `lakename`, `year4`, `daynum`, `depth`, `temperature_C`
 * Only complete cases (i.e., remove NAs)

5. Visualize the relationship among the two continuous variables with a scatter plot of temperature by depth. Add a smoothed line showing the linear model, and limit temperature values from 0 to 35 °C. Make this plot look pretty and easy to read.

```{r scatterplot}
#4
NTL_Chemphy_July <- NTL_Chemphy %>%
  filter( month (sampledate) == 7) %>%
  select(lakename, year4, daynum, depth, temperature_C) %>%
  na.omit()

#5
temperaturedepth_July <-
  ggplot(NTL_Chemphy_July, aes(x = depth, y = temperature_C)) +
  geom_point()+
  ylim(0,35) +
  geom_smooth(method = "lm")+
  labs(x="Depth (m)", y=expression ("Temperature ("^o*"C)"))

print(temperaturedepth_July)

```


6. Interpret the figure. What does it suggest with regards to the response of temperature to depth? Do the distribution of points suggest about anything about the linearity of this trend?

> Answer: The figure suggest that the temperature in July will decrease with increasing depth. The disribution of points can suggest a negative linear relationship between the two variables.


7. Perform a linear regression to test the relationship and display the results

```{r linear.regression}
#7
temperature.regression <- lm(data = NTL_Chemphy_July, temperature_C ~ depth)
summary(temperature.regression)

```


8. Interpret your model results in words. Include how much of the variability in temperature is explained by changes in depth, the degrees of freedom on which this finding is based, and the statistical significance of the result. Also mention how much temperature is predicted to change for every 1m change in depth. 

> Answer:
An increase of one meter of depth is associated with a decrease of 1.9 degrees of celsius in the lake. This coefficient is statistically different than 0 (p<0.001), so we reject the null hypothesis. R^2 is the percent of the variation in Y variable that is explained by the X variable. Our R^2 is 0.74, which is 74% of variations in temperature can be explained by depth. Residual Standard Error is measure of the quality of a linear regression fit, in which the actual temperature can deviate from the true regression line by approximately 3.8 degrees of celsius. Degrees of freedom are the number of data points that went into the estimation of the parameters used after taking into account these parameters (restriction). In our case, we had 9728 data points and two parameters (intercept and slope).

---

## Multiple regression
Let's tackle a similar question from a different approach. Here, we want to explore what might the best set of predictors for lake temperature in July across the monitoring period at the North Temperate Lakes LTER. 


9. Run an AIC to determine what set of explanatory variables (year4, daynum, depth) is best suited to predict temperature.

10. Run a multiple regression on the recommended set of variables. 

```{r temperature.model}
#9
NTL_AIC <- lm(data = NTL_Chemphy_July, temperature_C ~ depth + year4 +
              daynum)

step(NTL_AIC)

#10
NTL_model <- lm(data = NTL_Chemphy_July, temperature_C ~ depth + year4 +
              daynum)
summary(NTL_model)

```

11. What is the final set of explanatory variables that the AIC method suggests we use to predict temperature in our multiple regression? How much of the observed variance does this model explain? Is this an improvement over the model using only depth as the explanatory variable?

> Answer: year4, daynum and depth are suggested to predict temperature in our multiple regression. Our R^2 is 0.74, which is 74% of variations in temperature can be explained by those three variables. This does not improve too much over the model using only depth as the explanatory variable.



---
## Analysis of Variance

12. Now we want to see whether the different lakes have, on average, different temperatures in the month of July. Run an ANOVA test to complete this analysis. (No need to test assumptions of normality or similar variances.) Create two sets of models: one expressed as an ANOVA models and another expressed as a linear model (as done in our lessons).

```{r anova.model}
#12
NTL_Chemphy_July.anova <- aov(data = NTL_Chemphy_July, temperature_C ~ lakename)

summary(NTL_Chemphy_July.anova)

NTL_Chemphy_July.anova2 <- lm(data = NTL_Chemphy_July, temperature_C ~ lakename)

summary(NTL_Chemphy_July.anova2)

```

13. Is there a significant difference in mean temperature among the lakes? Report your findings. 

> Answer: Since the p-value in our ANOVA table is less than .001, we have sufficient evidence to reject the null hypothesis (H0: all means among lakes are same). Thus, mean temeprature is different among the lakes.



14. Create a graph that depicts temperature by depth, with a separate color for each lake. Add a geom_smooth (method = "lm", se = FALSE) for each lake. Make your points 50 % transparent. Adjust your y axis limits to go from 0 to 35 degrees. Clean up your graph to make it pretty. 

```{r scatterplot.2}
#14.
temperaturedepth_seperate_July <-
  ggplot(NTL_Chemphy_July, aes(x = depth, y = temperature_C, color=lakename)) +
  geom_point(alpha = 0.5)+
  ylim(0,35) +
  geom_smooth(method = "lm",  se = FALSE)+
  labs(x="Depth (m)", y=expression ("Temperature ("^o*"C)"))

print(temperaturedepth_seperate_July)

```

15. Use the Tukey's HSD test to determine which lakes have different means.

```{r tukey.test}
#15
TukeyHSD(NTL_Chemphy_July.anova)

```

16.From the findings above, which lakes have the same mean temperature, statistically speaking, as Peter Lake? Does any lake have a mean temperature that is statistically distinct from all the other lakes?

>Answer:Paul Lake and Ward Lake have same mean temeprature statistically. There is no lake have a mean temperature that is statistically distinct from all the other lakes
 

17. If we were just looking at Peter Lake and Paul Lake. What's another test we might explore to see whether they have distinct mean temperatures? 

>Answer: two sample t test



18. Wrangle the July data to include only records for Crampton Lake and Ward Lake. Run the two-sample T-test on these data to determine whether their July temperature are same or different. What does the test say? Are the mean temperatures for the lakes equal? Does that match you answer for part 16?

```{r t.test}
crampton_ward_July <- NTL_Chemphy_July %>%
  filter(lakename == "Crampton Lake"| lakename == "Ward Lake")

crampton_ward_July.twosample <- t.test(crampton_ward_July$temperature_C ~ crampton_ward_July$lakename)

crampton_ward_July.twosample 
```

>Answer: In this t test, the mean July temperature of Crampton Lake and Ward Lake are compared. The null hypothesis is there is no difference in the mean July temperature between two lakes (µA-µB=0). The alternative hypothesis is there is difference in the mean July temperature between two lakes (µA-µB not equal to 0). There is no evidence (t=1.11, df=200.37, p~0.26) to reject the null hypothesis, so it suggests that no difference in the mean July temperature between two lakes. This result match with the result in part 16.
